Clustering Beat-Chroma Patterns in a Large Music Database

### 2010-03-23 Paper/Paper.tex
fixed i.e.'s and e.g.'s

\begin{document}
%
\maketitle
%
\begin{abstract}
A musical style or genre implies a set of common conventions
and patterns combined and deployed in different ways to make
individual musical pieces; for instance, most would agree that
contemporary pop music is assembled from a relatively small
palette of harmonic and melodic patterns.  The purpose of this
paper is to use a database of tens of thousands of songs
in combination with a compact representation of melodic-harmonic
content, the beat-synchronous chromagram, and data-mining
tools (clustering) to attempt to explicitly catalog this palette --
at least within the limitations of the beat-chroma representation.
We use online $k$-means clustering to summarize 3.7 million
4-beat bars in a codebook of a few hundred prototypes.
By measuring how accurately such a quantized codebook
can reconstruct the original data, we can quantify the degree
of diversity (distortion as a function of codebook size) and
temporal structure (i.e. the advantage gained
by joint quantizing multiple frames) in this music.  The most
popular codewords themselves reveal the common chords
used in the music.  Finally, the quantized representation of
music can be used for music retrieval tasks such as artist
and genre classification, and identifying songs that are
similar in terms of their melodic-harmonic content.
\end{abstract}
%
\section{Introduction}\label{sec:introduction}
The availability of very large collections of music audio present
many interesting research opportunities.  Given millions of examples
from a single, broad class (e.g., contemporary commercial pop music),
can we infer anything about the underlying structure and common
features of this class?  This paper describes our work in this direction.

What are the common features of pop music?  There are conventions
of subject matter, instrumentation, form, rhythm, harmony, and melody, among
others.  Our interest here is in the tonal content of the music -- i.e. the
harmony and melody.  As a computationally convenient proxy for a richer
description of the tonal content of audio, we use the popular Chroma
representation, which collapses an acoustic spectrum into a 12-dimensional
description, with one bin for each semitone of the western musical octave.
In addition, we simplify the time axis of our representation to take advantage
of the strong beat present in most pop music, and record just one
chroma vector per beat.  This beat-synchronous chromagram
representation represents a typical music track
in a few thousand values, yet when resynthesized back into audio via
modulation of octave-invariant ``Shepard tones'', the melody and
chord sequences of the original music typically remain recognizable
\cite{Ellis2007a}.  To the extent, then, that beat-chroma
representations preserve tonal structure, they
are an interesting domain in which to search for patterns -- rich enough
to generate musically-relevant results, but simplified enough to
abstract away some other aspects of the original audio, such as
instrumentation and other stylistic details.

@@

Note that we do not claim that any of these informations (segments, beats, bars)
are perfectly accurate (e.g. in \cite{Barrington2009a} authors obtain a better
song segmentation than Echonest). Practise showed us that they are reasonable, 
and the size of the data set should make up for the imperfections or noise.
We also believe that patches sizes based on a number of beats or bars are more
meaningful than an arbitrary time length. More on this in the experiments
(Section \ref{sec:experiments} TO BE CHECKED).

@@

We use the codebook to encode our testing set (see Section \ref{sec:testdata}).
Each pattern is encoded with only one code. We can measure the average
distance between a pattern and its encoding. We can also measure the use
of the different codes, i.e. the size of the cluster around a code.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{codes_lle}
\end{center}
\caption{\small{
Locally linear embedding (LLE) visualization of the codebook.
LLE tries to keep near neighbors (here according to euclidean distance)
in the same $2$D neighborhood. Shown patterns were randomly selected
in each neighborhood.
}}
\label{fig:lle}
\end{figure}

@@

\subsection{Cluster}
Once we understand the kind of codebook we create, we naturally want
to visualize the clusters, i.e. the pattersn from actual song that
surround a given centroid. Unsurprisingly, real patterns tend to be
noisier than codes. However, centroids seem to capture important aspects
of the patterns. We could compare the centroids with the result of
a lowpass filtering.

>>>>

\begin{document}
%
\maketitle
%
\begin{abstract}
A musical style or genre implies a set of common conventions
and patterns combined and deployed in different ways to make
individual musical pieces; for instance, most would agree that
contemporary pop music is assembled from a relatively small
palette of harmonic and melodic patterns.  The purpose of this
paper is to use a database of tens of thousands of songs
in combination with a compact representation of melodic-harmonic
content, the beat-synchronous chromagram, and data-mining
tools (clustering) to attempt to explicitly catalog this palette --
at least within the limitations of the beat-chroma representation.
We use online $k$-means clustering to summarize 3.7 million
4-beat bars in a codebook of a few hundred prototypes.
By measuring how accurately such a quantized codebook
can reconstruct the original data, we can quantify the degree
of diversity (distortion as a function of codebook size) and
temporal structure (\ie the advantage gained
by joint quantizing multiple frames) in this music.  The most
popular codewords themselves reveal the common chords
used in the music.  Finally, the quantized representation of
music can be used for music retrieval tasks such as artist
and genre classification, and identifying songs that are
similar in terms of their melodic-harmonic content.
\end{abstract}
%
\section{Introduction}\label{sec:introduction}
The availability of very large collections of music audio present
many interesting research opportunities.  Given millions of examples
from a single, broad class (\eg, contemporary commercial pop music),
can we infer anything about the underlying structure and common
features of this class?  This paper describes our work in this direction.

What are the common features of pop music?  There are conventions
of subject matter, instrumentation, form, rhythm, harmony, and melody, among
others.  Our interest here is in the tonal content of the music -- \ie the
harmony and melody.  As a computationally convenient proxy for a richer
description of the tonal content of audio, we use the popular Chroma
representation, which collapses an acoustic spectrum into a 12-dimensional
description, with one bin for each semitone of the western musical octave.
In addition, we simplify the time axis of our representation to take advantage
of the strong beat present in most pop music, and record just one
chroma vector per beat.  This beat-synchronous chromagram
representation represents a typical music track
in a few thousand values, yet when resynthesized back into audio via
modulation of octave-invariant ``Shepard tones'', the melody and
chord sequences of the original music typically remain recognizable
\cite{Ellis2007a}.  To the extent, then, that beat-chroma
representations preserve tonal structure, they
are an interesting domain in which to search for patterns -- rich enough
to generate musically-relevant results, but simplified enough to
abstract away some other aspects of the original audio, such as
instrumentation and other stylistic details.

@@

Note that we do not claim that any of these informations (segments, beats, bars)
are perfectly accurate (\eg in \cite{Barrington2009a} authors obtain a better
song segmentation than Echonest). Practise showed us that they are reasonable, 
and the size of the data set should make up for the imperfections or noise.
We also believe that patches sizes based on a number of beats or bars are more
meaningful than an arbitrary time length. More on this in the experiments
(Section \ref{sec:experiments} TO BE CHECKED).

@@

We use the codebook to encode our testing set (see Section \ref{sec:testdata}).
Each pattern is encoded with only one code. We can measure the average
distance between a pattern and its encoding. We can also measure the use
of the different codes, \ie the size of the cluster around a code.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.9\columnwidth]{codes_lle}
\end{center}
\caption{\small{
Locally linear embedding (LLE) visualization of the codebook.
LLE tries to keep near neighbors (here according to euclidean distance)
in the same $2$D neighborhood. Shown patterns were randomly selected
in each neighborhood.
}
ronw: I think this figure would be better if it only included the
bottom subplot.
}
\label{fig:lle}
\end{figure}

@@

\subsection{Cluster}
Once we understand the kind of codebook we create, we naturally want
to visualize the clusters, \ie the pattersn from actual song that
surround a given centroid. Unsurprisingly, real patterns tend to be
noisier than codes. However, centroids seem to capture important aspects
of the patterns. We could compare the centroids with the result of
a lowpass filtering.
